---
title: "p8105_hw3_yj2687"
author: "Yijia Jiang"
date: "2022-10-07"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```




## Problem 1 (Instacart dataset)
```{r,warning = FALSE,message=FALSE}
data("instacart")

instacart_df = 
  instacart %>% 
  as_tibble(instacart)

instacart_df %>% 
  count(aisle) %>% 
  arrange(desc(n))

instacart_df %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()

instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

&nbsp;



## Problem 2 (Accelerometer dataset)
```{r,warning = FALSE,message=FALSE}
# Tidy the dataset
accel_df = read.csv("./p8105_hw3_data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(activity_1:activity_1440, names_to = "minutes_in_a_day", 
               names_prefix = "activity_", values_to = "activity_count") %>% 
  mutate(weekday_vs_weekend = case_when(
    day == "Monday"  ~ "Weekday",
    day == "Tuesday"  ~ "Weekday",
    day == "Wednesday"  ~ "Weekday",
    day == "Thursday"  ~ "Weekday",
    day == "Friday"  ~ "Weekday",
    day == "Sunday"  ~ "Weekend",
    day == "Saturday"  ~ "Weekend"
  )) %>% 
  select(week, day_id, day, weekday_vs_weekend, everything()) %>% 
  mutate_if(is.double, as.integer) %>%
  mutate(minutes_in_a_day = as.integer(minutes_in_a_day))
```

* There are `r nrow(accel_df)` observations in the resulting tidy dataset, including `r ncol(accel_df)` variables, namely `r names(accel_df)`.


```{r}
# Create a table showing total activity for each day by aggregating across minutes
accel_df$day = factor(accel_df$day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))

accel_df %>% 
  group_by(day) %>% 
  summarize(total = sum(activity_count)) %>% 
  knitr::kable(align = "l", format = "pipe", col.names = c("Day","Total"))

```

* It is difficult to identify the apparent trends across days according to this table, while we can see the total activity on Saturdays is much lower than that on other days and Friday has the relatively highest total activity.


```{r}
# Make a single-panel plot showing the 24-hour activity time courses for each day and use color to indicate day of the week. 
accel_df %>% 
  group_by(day, minutes_in_a_day) %>%
  rename(Day = day) %>%
  ggplot(aes(minutes_in_a_day, activity_count, color = Day)) +
  geom_point(alpha = 0.5) +
  scale_x_continuous(
    breaks = c(0, 180, 360, 540, 720, 900, 1080, 1260, 1440),
    labels = c("12AM", "3AM", "6AM", "9AM", "12PM", "3PM", "6PM", "9PM", "11:59PM")) + 
  labs(
    title = "24-hour activity time courses for each day",
    x = "Time",
    y = "Activity Count") +
  theme(legend.position = "right")
```

* We can see that most of the activity count for every minute throughout the day is usually below 2500. The activity count at noon (12PM) and evenings (9PM) tends to peak compared to the other time of the day, and the activity count tends to be very low between the period [12AM,6AM] across days as people normally sleep during this time. In terms of point dispersion, there are more activity counts per minute on Friday and Wednesday.

&nbsp;



## Problem 3 (NY NOAA dataset)
```{r,warning = FALSE,message=FALSE}
# Tidy the dataset
noaa_df = ny_noaa %>% 
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  arrange(year, month) %>%
  mutate(prcp = prcp/10,
         tmax = (as.double(tmax)/10),
         tmin = (as.double(tmin)/10)) 

# Count the most commonly values for snowfall
snow_obs = noaa_df %>%
  count(snow) %>% 
  arrange(desc(n))
```

* This data collected weather data from all New York state weather stations between `r noaa_df %>% pull(year) %>% min()` and `r noaa_df %>% pull(year) %>% max()`.
* There are `r nrow(noaa_df)` observations in the resulting tidied dataset, including `r ncol(noaa_df)` variables, namely `r names(noaa_df)`.
* The `id` is a `r class(noaa_df$id)` variable, the `year`, `month`, `day` are `r class(noaa_df$year)` variables, and `prcp`, `snow`, `snwd`, `tmax`, `tmin` are `r class(noaa_df$prcp)` variables.
* Interestingly, before tidy procedure, the variables `tmax` and `tmin` were defined as a character, the `snow` variable has a negative value of -13mm.
* There exist `r sum(is.na(noaa_df$prcp))` missing values in `prcp`, `r sum(is.na(noaa_df$snow))` missing values in `snow`, `r sum(is.na(noaa_df$snwd))` missing values in `snwd`, `r sum(is.na(noaa_df$tmax))` missing values in `tmax` and `r sum(is.na(noaa_df$tmin))` missing values in `tmin`.
* For snowfall, the most commonly observed value is `r snow_obs[1,1]` mm.


```{r, warning = FALSE,message=FALSE}
# Make a two-panel plot showing the average max temperature in January and in July in each station across years.
jan_vs_july = noaa_df %>%
  filter(!is.na(tmax), month %in% c("1", "7")) %>%
  mutate(month = recode(month, "01" = "January", "07" = "July")) %>%
  group_by(id, year, month) %>%
  summarise(average_tmax = mean(tmax))

# Scatterplot
jan_vs_july %>% 
  ggplot( aes(x = year, y = average_tmax)) +
  geom_point(alpha = .5) +
  geom_smooth() +
  facet_grid(. ~ month) +
  labs(
    title = "Average max temperature in January and in July in each station across years",
    x = "Year",
    y = "Average Maximum Temperature (C)",
    caption = "Data from the noaa package") 
```

* This two-panel scatter plot compares the average maximum temperature in January and in July in each station across years. We can observe that the average highest temperature in January is much lower than that in July across 30 years. This is because January is still winter and July is summer for New York. Additionally, the fluctuation of maximum temperature in January is larger than that in July, in a wavy pattern, first decreasing year by year and then beginning to rise year by year. Relatively, the maximum temperatures vary less in July. For both months, there are outliers present across the years, the average max temperature in 1988 July is abnormally low. 


```{r, warning = FALSE,message=FALSE}
# Make a plot showing tmax vs tmin for the full dataset
# Hex plot
p1= noaa_df %>%
  filter(!is.na(tmin), !is.na(tmax)) %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() +
  labs(
    title = "Max and min temperature comparison",
    x = "Max temperature (C)",
    y = "Min temperature (C)") +
  theme(legend.title = element_text(size = 9), 
        legend.text  = element_text(size = 5))

# Make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year
p2 = noaa_df %>%
  filter(snow > 0 & snow < 100) %>%
  group_by(year) %>%
  ggplot(aes(x = factor(year), y = snow)) +
  geom_boxplot() +
  labs(
    title = "The distribution of snowfall",
    x = "Year",
    y = "Snowfall (mm)",
    caption = "Data from the noaa package") +
  theme(axis.text.x = element_text(size = 8, angle = 90, hjust = 0)) +
  guides(color = guide_legend(override.aes = list(size = 5)))

p2 = noaa_df %>% 
  filter(snow > 0 & snow < 100) %>%
  ggplot(aes(x = snow, y = as.factor(year))) +
  geom_density_ridges(scale = 0.8) +
  labs(
    title = "The distribution of snowfall",
    x = "Snowfall (mm)",
    y = "Year",
    caption = "Data from the noaa package") 

# Two-panel layout
p1+p2
```

* A two-panel plot is created by combining the hex plot and the density ridge plot, showing the comparison between the maximum temperature and the minimum temperature, and the distribution of snowfall values greater than 0 and less than 100 separately by year. From the hex plot, we observe that the temperature is around 0 to 25 degrees for most of the time. From the density ridge plot, we observe similar pattern of snowfall distribution across years by excluding the values less than 0 and more than 100, which is right skewed. Most of the snowfall has the value ranging from 10 mm to 30 mm, with some falling as much as 50 mm. 

